### 2019-01-19 figures/general_information/elife_general_information.rmd

---
title: "elife_not_anon Dataset—General Information"
author: "Dakota Murray"
output: html_notebook
---

## Script settings
```{r}
path_to_data_elife <- "https://www.dropbox.com/s/dmvmzz57m7t4xtq/non_anonymous_data.csv?dl=1"
path_to_data_people <- "https://www.dropbox.com/s/c8r7goz9pdsr4hd/elife_people.csv?dl=1"
```

```{r message=FALSE, warning=FALSE}
# Load required packages
library(tidyverse)
library(stargazer)
```

## Load the data
```{r message=FALSE, warning=FALSE}
elife_not_anon <- read_csv(path_to_data_elife)
people <- read_csv(path_to_data_people)
head(elife_not_anon)
```

Lets calculate some basic things, before we anonymize the data, like number of unique co-authors, etc.

## Initial Submissions and General Info

### Number of initial submissions
```{r}
dim(elife_not_anon)[1]
```

### Distribution of manuscript types
```{r}
table(elife_not_anon$type.x)
```

### Distribution of manuscript types, as %
```{r}
prop.table(table(elife_not_anon$type.x))
```

### Review outcome of initial submissions to elife_not_anon
```{r}
outcomes <- elife_not_anon %>%
  mutate(outcome = ifelse(!initial_decision_made, "No Decision Yet", ifelse(encouraged, "Encouraged", "Rejected")))

table(outcomes$outcome)
```

## Proportion of initial submissions with decisions which were accepted/rejected
```{r}
with_decision <- outcomes %>% filter(outcome != "No Decision Yet")

prop.table(table(with_decision$outcome))
```

### Gender dsitirbution of distinct corresponding authors of initial submissions to elife_not_anon 
```{r}
# get distint corresondign authros on first submissions
ca_authors <- elife_not_anon %>% 
  mutate(ca_name = tolower(ca_name)) %>%
  group_by(ca_name, ca_country) %>%
  filter(row_number() == 1)

table(ca_authors$ca_gender)
```

### National distribution of distinct corresonding authors of initial submissions
```{r}
t <- table(ca_authors$ca_country)
t[order(-t)]
```

### Gender distibution of distinct senior editors of initial submissions 
```{r}
# get distint corresondign authros on first submissions
se_authors <- elife_not_anon %>% 
  mutate(se_name = tolower(se_name)) %>%
  group_by(se_name, se_country) %>%
  filter(row_number() == 1)

table(se_authors$se_gender)
```

### National distribution of distinct senior editors of initial submissions
```{r}
t <- table(se_authors$se_country)
t[order(-t)]
```

## Full Submissions

### Total number of full submissions
```{r}
full <- elife_not_anon %>%
  filter(has_full_submission)

dim(full)[1]
```

###Final decision outcomes for full submissions
```{r}
outcomes <- full %>%
  mutate(outcome = ifelse(!final_decision_made, "No Decision Yet", ifelse(accepted, "Accepted", "Rejected")))

table(outcomes$outcome)
```

### First full decision outcomes for full submissions
```{r}
table(ifelse(is.na(full$full_decision), "No Decision", full$full_decision))
prop.table(table(ifelse(is.na(full$full_decision), "No Decision", full$full_decision)))
```

###First revision review outcomes for full submissions
```{r}
rev1 <- full %>% filter(!is.na(rev1_submission_date))
table(ifelse(is.na(rev1$rev1_decision), "No Decision", rev1$rev1_decision))
```

###Second revision review outcomes for full submissions
```{r}
rev2 <- full %>% filter(!is.na(rev2_submission_date))
table(ifelse(is.na(rev2$rev2_decision), "No Decision", rev2$rev2_decision))
```

###Number of distinct corresponding authors in full submissions at elife_not_anon
```{r}
ca_authors <- full %>% 
  filter(!is.na(ca_name)) %>%
  mutate(ca_name = tolower(ca_name)) %>%
  group_by(ca_name, ca_country) %>%
  filter(row_number() == 1)

dim(ca_authors)[1]
```

###Gender distirbution of distinct correpsonding authors on full submissions at elife_not_anon
```{r}
table(ca_authors$ca_gender)
```

###Number of distinct first authors on full submissions at elife_not_anon
```{r}
fa_authors <- full %>% 
  filter(!is.na(fa_name)) %>%
  mutate(fa_name = tolower(fa_name)) %>%
  group_by(fa_name, fa_country) %>%
  filter(row_number() == 1)

dim(fa_authors)[1]
```

###Gender distirbution of distinct first authors on full submissions at elife_not_anon
```{r}
table(fa_authors$fa_gender)
prop.table(table(fa_authors$fa_gender))
```

###Gender distirbution of distinct gender-identified first authors on full submissions at elife_not_anon
```{r}
identified <- fa_authors %>% filter(fa_gender %in% c("M", "F"))
table(identified$fa_gender)
prop.table(table(identified$fa_gender))
```

###Number of distinct last authors on full submissions at elife_not_anon
```{r}
la_authors <- full %>% 
  filter(!is.na(la_name)) %>%
  mutate(la_name = tolower(la_name)) %>%
  group_by(la_name, la_country) %>%
  filter(row_number() == 1)

dim(la_authors)[1]
```

### Gender distirbution of distinct last authors on full submissions at elife_not_anon
```{r}
table(la_authors$la_gender)
```

### Gender distirbution of distinct gender-identified first authors on full submissions at elife_not_anon
```{r}
identified <- la_authors %>% filter(la_gender %in% c("M", "F"))
table(identified$la_gender)
prop.table(table(identified$la_gender))
```

### Gender distirbution of distinct last authors on full submissions at elife_not_anon, as %
```{r}
prop.table(table(la_authors$la_gender))
```

### Number of distinct gatekeepers involved in processing full submissions at elife_not_anon, includes senior editors, reviewing editors, and invited peer reviewers
```{r}
gatekeepers <- people %>% filter(type == "Gatekeeper")
dim(gatekeepers)[1]
```

### Gender distirbution of distinct gatekeepers involved in processing full submissions at elife_not_anon, includes senior editors, reviewing editors, and invited peer reviewers
```{r}
table(gatekeepers$gender)
prop.table(table(gatekeepers$gender))
```

###Gender distribution of distinct gatekeepers invovled in the processing of full submissions submissions at elife_not_anon, but now for only gender-identified individuals
```{r}
identified <- gatekeepers %>% filter(gender %in% c("M", "F"))
table(identified$gender)
prop.table(table(identified$gender))
```

### Continent distirbution of distinct gatekeepers involved in processing full submissions at elife_not_anon, includes senior editors, reviewing editors, and invited peer reviewers
```{r}
table(gatekeepers$continent)
prop.table(table(gatekeepers$continent))
```

###Proportion of full submissions for which the correposnding author was also the first author
```{r}
sum(full$ca_name == full$fa_name, na.rm = T) / dim(full)[1]
```

###Proportion of full submissions for which the correposnding author was also the last author
```{r}
sum(tolower(full$ca_name) == tolower(full$la_name), na.rm = T) / dim(full)[1]
```

###Number of full submissions that were appealed
```{r}
table(full$appealed_any_stage)
```

### Average number of revisions before final acceptance of full submissions at elife_not_anon
```{r}
# In this case, "0" means 
f <- full %>% filter(accepted)
mean(f$num_revisions, na.rm = T)
```

###Proportion of full submissions for which the correpsonding auhtor is from the same country as the first author
```{r}
sum(full$ca_country == full$fa_country, na.rm = T) / dim(full)[1]
```

###Proportion of full submissions for which the correpsonding auhtor is from the same country as the last author
```{r}
sum(full$ca_country == full$la_country, na.rm = T) / dim(full)[1]
```

###Number of unique name/role combinations that appear in the data
```{r}
dim(people)[1]
```

###Number of thees combinations for which a gender was assigned
```{r}
sum(people$gender != "UNK")
sum(people$gender != "UNK") / dim(people)[1]

```

###Gender distribution of all name/role combinations that appear in the dataset
```{r}
table(people$gender)
prop.table(table(people$gender))
```

```{r}
full <- elife_not_anon %>% 
  filter(has_full_submission & final_decision_made) %>%
  filter(la_gender %in% c("M", "F")) %>%
  filter(composition != "Uncertain")

summary(glm(accepted ~ la_gender * composition, data = full, family = "binomial"))
```

## Table SI.2

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
gatekeepers <- subset(people, type == "Gatekeeper")

# Get a single list of all the peer reviewers, which presently are represented as separate variables on each row. This method
# is ugly and can be improved, but it is simple and works, so I won't bother changing it
reviewer1 <- elife_not_anon[, c("reviewer1_name", "reviewer1_gender", "reviewer1_institution", "reviewer1_country")]
reviewer1 <- reviewer1[with(reviewer1, !is.na(reviewer1_name)), ]
colnames(reviewer1) <- c("reviewer_name", "reviewer_gender", "reviewer_institution", "reviewer_country")

reviewer2 <- elife_not_anon[, c("reviewer2_name", "reviewer2_gender", "reviewer2_institution", "reviewer2_country")]
reviewer2 <- reviewer2[with(reviewer2, !is.na(reviewer2_name)), ]
colnames(reviewer2) <- c("reviewer_name", "reviewer_gender", "reviewer_institution", "reviewer_country")

reviewer3 <- elife_not_anon[, c("reviewer3_name", "reviewer3_gender", "reviewer3_institution", "reviewer3_country")]
reviewer3 <- reviewer2[with(reviewer3, !is.na(reviewer3_name)), ]
colnames(reviewer3) <- c("reviewer_name", "reviewer_gender", "reviewer_institution", "reviewer_country")

reviewer4 <- elife_not_anon[, c("reviewer4_name", "reviewer4_gender", "reviewer4_institution", "reviewer4_country")]
reviewer4 <- reviewer2[with(reviewer4, !is.na(reviewer4_name)), ]
colnames(reviewer4) <- c("reviewer_name", "reviewer_gender", "reviewer_institution", "reviewer_country")

# Combine all of these reviewer datasets into one
reviewers <- rbind(reviewer1, reviewer2, reviewer3, reviewer4)

bres <- elife_not_anon[, c("bre_gender", "bre_name", "bre_country")]
peer_reviewers <- reviewers[, c("reviewer_gender", "reviewer_name", "reviewer_country")]
editors <- elife_not_anon[, c("se_gender", "se_name", "se_country")]
names(editors) <- c("gender", "name", "country")
names(bres) <- c("gender", "name", "country")
names(peer_reviewers) <- c("gender", "name", "country")

editors <- subset(editors, !duplicated(name, country))

reviewers <- subset(reviewers, !duplicated(reviewer_name, reviewer_country))
bres <- subset(bres, !duplicated(name, country))

t1 <- data.frame(table(editors$country), prop.table(table(editors$country)))
t2 <- data.frame(table(reviewers$reviewer_country), prop.table(table(reviewers$reviewer_country)))
t3 <- data.frame(table(bres$country), prop.table(table(bres$country)))

all_data <- t2 %>% 
  left_join(t3, by = "Var1") %>%
  left_join(t1, by = "Var1") %>%
  select(-Var1.1.x, -Var1.1.y, -Var1.1)

names(all_data) <- c("country", "rev.freq", "rev.prop", "bre.freq", "bre.prop", "se.freq", "se.prop")

ord <- with(all_data, order(-se.freq, -bre.freq))

all_data <- all_data[ord, ]

stargazer(all_data, type = "html")
```

###Gender distribution of gender-identified correpsonding authors of full submissions
```{r}
identified <- full %>% filter(ca_gender != "UNK")
table(identified$ca_gender)
prop.table(table(identified$ca_gender))
```

###Gender distribution of gender-identified first authors of full submissions
```{r}
identified <- full %>% filter(fa_gender != "UNK")
table(identified$fa_gender)
prop.table(table(identified$fa_gender))
```

###Gender distribution of gender-identified last authors of full submissions
```{r}
identified <- full %>% filter(la_gender != "UNK")
table(identified$la_gender)
prop.table(table(identified$la_gender))
```