### 2019-01-19 figures/selectivity_over_time/selectivity_over_time.rmd

---
title: "Submissions and selectivity of *eLife* over time."
author: "Dakota Murray"
output: html_notebook
---

### Set parameters for the name and extension of the figure
```{r}
figure_name <- "fig_2"
figure_extension <- ".png"

path_to_data <- "https://www.dropbox.com/s/5doa2khbrs7nuyx/formatted_elife.csv?dl=1"
```

```{r message=FALSE, warning=FALSE}
# Load required packages
library(tidyverse)
library(grid)
library(gridExtra)

# setup theme
theme_dakota <- function () { 
    theme_bw(base_size=10, base_family="Times") %+replace% 
        theme(
            strip.background = element_blank(),
            strip.text = element_text(face = "bold"),
            plot.title = element_text(size = 11, face = "bold"),
            axis.text.x = element_text(size = 9, margin = margin(t = 5, b = 5)),
            legend.title = element_blank(),
            legend.position = "bottom"
        )
}
```

### Load the required data
```{r message=FALSE, warning=FALSE}
elife <- read_csv(path_to_data)
head(elife)
```

### Plot the figure

Note: Figure 1 wa smade in an external program, *RawGraphics*, and so does not appear in this notebook.

> Submissions and selectivity of eLife over time. Left: Yearly count of initial submissions, encouraged initial submissions, and accepted full submissions to eLife between 2012 and 2016; Right: rate of initial submissions encouraged (Encourage %), rate of full submissions accepted (% Full accepted) and rate of initial submissions accepted (Overall accept %) between 2012 and 2016. Submissions during the year of 2017 were excluded because we do not have sufficient data for full life-cycle of these manuscripts.

```{r}
d1 <- elife %>%
  group_by(submission_year) %>%
  summarize(
        paper_count = n(),
        accept_count = sum(accepted == TRUE, na.rm = T),
        encourage_count = sum(encouraged == TRUE, na.rm = T)
  ) %>%
  gather(measure_key, value = measurement, paper_count, encourage_count, accept_count) %>%
  mutate(measure_key = factor(measure_key, levels = c("paper_count", "encourage_count", "accept_count"))) %>%
  filter(submission_year != "2017")

p1 <- d1 %>%
  ggplot(aes(x = submission_year, y = measurement, linetype = measure_key, color = measure_key, shape = measure_key)) +
  geom_line() +
  geom_point(size = 2) +
  theme_dakota() +
  labs(
    x = "Submission Year",
    y = "# Manuscripts"
  ) +
  scale_color_manual(labels = c("accept_count" = "# Accepted", "encourage_count" = "# Encouraged", "paper_count" = "# Submitted"), values = c("black", "blue", "red")) +
  scale_shape_discrete(labels = c("accept_count" = "# Accepted", "encourage_count" = "# Encouraged", "paper_count" = "# Submitted")) +
  scale_linetype_discrete(labels = c("accept_count" = "# Accepted", "encourage_count" = "# Encouraged", "paper_count" = "# Submitted"))

d2 <- elife %>%
  group_by(submission_year) %>%
  summarize(
        paper_count = n(),
        accept_count = sum(accepted == TRUE, na.rm = T),
        encourage_count = sum(encouraged == TRUE, na.rm = T),
        encourage_rate = encourage_count / paper_count,
        accept_rate = accept_count / paper_count,
        encourage_accept_rate = accept_count / sum(!is.na(full_decision_made) & full_decision_made)
  ) %>%
  gather(measure_key, value = measurement, encourage_rate, accept_rate, encourage_accept_rate) %>%
  mutate(
    measure_key = factor(measure_key, levels = c("encourage_rate", "encourage_accept_rate", "accept_rate"))
  ) %>%
  filter(submission_year < 2017)

p2 <- d2 %>%
  ggplot(aes(x = submission_year, y = measurement, linetype = measure_key, color = measure_key, shape = measure_key)) +
  geom_line() +
  geom_point(size = 2) +
  ylim(0, 1) +
  theme_dakota() +
  labs(
    x = "Submission Year",
    y = "%"
  ) +
  scale_color_manual(labels = c("accept_rate" = "Overall accept %", "encourage_rate" = "Encourage %", "encourage_accept_rate" = "% Full accepted"), 
                     values = c("blue", "purple", "red")) +
  scale_shape_manual(labels = c("accept_rate" = "Overall accept %", "encourage_rate" = "Encourage %", "encourage_accept_rate" = "% Full accepted"),
                     values = c(17, 15, 18)) +
  scale_linetype_manual(labels = c("accept_rate" = "Overall accept %", "encourage_rate" = "Encourage %", "encourage_accept_rate" = "% Full accepted"), 
                          values = c("dotted", "dashed", "dotdash"))

selectivity_over_time <- grid.arrange(p1, p2, ncol = 2)
selectivity_over_time
```

### Save the figure and associated data
```{r}
name <- paste0("~/Dropbox/eLife/elife-analysis/figures/main/", figure_name, figure_extension)
ggsave(name, plot = selectivity_over_time, height = 4, width = 7.5)

name_csv1 <- paste0("~/Dropbox/eLife/elife-analysis/formatted_data/", figure_name, "_leftpanel_data.csv")
name_csv2 <- paste0("~/Dropbox/eLife/elife-analysis/formatted_data/", figure_name, "_rightpanel_data.csv")
write.csv(d1, name_csv1)
write.csv(d2, name_csv2)
```

### Data to construct figure 2, right panel, including enocurage, overall accept, and accept rates
```{r}
d2 %>% 
  spread(measure_key, measurement) %>%
  rename(
    overall_accept_rate = accept_rate,
    accept_rate = encourage_accept_rate
    )
```