### 2019-01-21 figures/appeals/gender_and_appeals.rmd
notebook to construct a figures about gender and appeals

---
title: "Gender and Appeals"
author: "Dakota Murray"
output: html_notebook
---

### Set parameters for the name and extension of the figure
```{r}
figure_name <- "fig_si7"
figure_extension <- ".png"

path_to_data <- "https://www.dropbox.com/s/97im2cayb1trlax/formatted_elife.csv?dl=1"
```

```{r message=FALSE, warning=FALSE}
# Load required packages
library(tidyverse)
library(grid)
library(gridExtra)

# setup theme
theme_dakota <- function () { 
    theme_bw(base_size=10, base_family="Times") %+replace% 
        theme(
            strip.background = element_blank(),
            strip.text = element_text(face = "bold"),
            plot.title = element_text(size = 11, face = "bold"),
            axis.text.x = element_text(size = 9, margin = margin(t = 5, b = 5)),
            legend.title = element_blank(),
            legend.position = "bottom"
        )
}
```

### Load the required data
```{r message=FALSE, warning=FALSE}
elife <- read_csv(path_to_data)

full_submissions <- elife %>% filter(full_decision_made)
full_submissions$included <- with(full_submissions, la_continent != "Antarctica" & fa_continent != "Antarctica" & !is.na(la_gender) & la_gender != "UNK" & composition != "Uncertain" & !is.na(la_reviewer_dist) & !is.na(la_reviewer_dist_zero) & !is.na(la_institution_rank) & !is.na(la_continent) & !is.na(fa_gender))

head(full_submissions)
```

```{r}
p_senior <- full_submissions %>%
  filter(included & !is.na(se_gender) & !is.na(la_gender)) %>%
  mutate(
    se_gender = factor(se_gender, labels = c("F" = "Female Senior Editor", "M" = "Male Senior Editor"))
  ) %>%
  group_by(la_gender, se_gender) %>%
  summarize(
    num = n(),
    num_appealed = sum(appealed_any_stage),
    appealed = sum(appealed_any_stage) / n(),
    ci = 1.96 * sqrt(appealed * (1 - appealed) / num)
  ) %>%
  ggplot(aes(x = la_gender, y = appealed, color = la_gender)) + 
  geom_point(size = 16, alpha = 0.5) +
  #geom_segment(aes(xend = la_gender, y = 0, yend = appealed - 0.0135)) +
  geom_errorbar(aes(ymin = appealed - ci, ymax = appealed + ci), width = 0.3) +
  facet_wrap(~se_gender) +
  guides(color = F, xend = F) +
  geom_label( aes(label = num_appealed)) +
  ylim(0, 0.15) +
  theme_bw() +
  labs(x = "Last Author Gender",
       y = "% of full submissions appealed",
       title = "Full Submissions appealed by gender of last author and senior editor")

p_reviewing <- full_submissions %>%
  filter(included & !is.na(bre_gender) & bre_gender != "UNK" & !is.na(la_gender)) %>%
  mutate(
    bre_gender = factor(bre_gender, labels = c("F" = "Female Reviewing Editor", "M" = "Male Reviewing Editor"))
  ) %>%
  group_by(la_gender, bre_gender) %>%
  summarize(
    num = n(),
    num_appealed = sum(appealed_any_stage),
    appealed = sum(appealed_any_stage) / n(),
    ci = 1.96 * sqrt(appealed * (1 - appealed) / num)
  ) %>%
  ggplot(aes(x = la_gender, y = appealed, color = la_gender)) + 
  geom_point(size = 16, alpha = 0.5) +
  geom_errorbar(aes(ymin = appealed - ci, ymax = appealed + ci), width = 0.3) +
  facet_wrap(~bre_gender) +
  guides(color = F, xend = F) +
  geom_label( aes(label = num_appealed)) +
  ylim(0, 0.20) +
  theme_bw() +
  labs(x = "Last Author Gender",
       y = "% of full submissions appealed",
       title = "Full Submissions appealed by gender of last author and reviewing editor")

g_appealed <- grid.arrange(p_senior, p_reviewing)
```

### Save the file
```{r}
name <- paste0("~/Dropbox/eLife/elife-analysis/figures/supp/", figure_name, figure_extension)
ggsave(name, plot = g_appealed, height = 7, width = 7)
```