### 2019-02-19 figures/selectivity_over_time/country_composition_shift.rmd

---
title: "Country Submissions Over Time"
author: "Dakota Murray"
output: html_notebook
---

### Set parameters for the name and extension of the figure
```{r}
figure_name <- "fig_si9"
figure_extension <- ".pdf"

path_to_data <- "https://www.dropbox.com/s/97im2cayb1trlax/formatted_elife.csv?dl=1"
```

```{r message=FALSE, warning=FALSE}
# Load required packages
library(tidyverse)
library(grid)
library(gridExtra)

# setup theme
theme_dakota <- function () { 
    theme_bw(base_size=10, base_family="Times") %+replace% 
        theme(
            strip.background = element_blank(),
            strip.text = element_text(face = "bold"),
            plot.title = element_text(size = 11, face = "bold"),
            axis.text.x = element_text(size = 9, margin = margin(t = 5, b = 5)),
            legend.title = element_blank(),
            legend.position = "bottom"
        )
}

# converts a provided p-value into a representation that is easier to visualize
sig2ast <- function(p) { ifelse(p <= 0.0001, "****", 
                                ifelse(p <= 0.001, "***", 
                                       ifelse(p <= 0.01, "**", 
                                              ifelse(p <= 0.05, "*", 
                                                     ifelse(p < 0.1, ".", "ns")
                                              )
                                       )
                                )
                          )
}

# I copied the below function from a stack overflow question. It should capitalize the first letter of each work in a string
# https://stackoverflow.com/questions/6364783/capitalize-the-first-letter-of-both-words-in-a-two-word-string
simpleCap <- function(x) {
  s <- strsplit(x, " ")[[1]]
  paste(toupper(substring(s, 1,1)), substring(s, 2),
      sep="", collapse=" ")
}
```

### Load the required data
```{r message=FALSE, warning=FALSE}
elife <- read_csv(path_to_data)
head(elife)
```

```{r}
country_over_time <- elife %>%
  #mutate(submission_year = submission_year + 2012) %>%
  filter(submission_year <= 2016 & !is.na(ca_continent)) %>%
  group_by(submission_year, ca_continent) %>%
  summarize(
    count = n()
  ) %>%
  ggplot(aes(x = submission_year, y = count, color = ca_continent, shape = ca_continent)) +
  geom_line() +
  geom_point() +
  theme_dakota() +
  labs(x = "Year",
       y = "# Submissions")

country_over_time
```

```{r}
country_over_time <- elife %>%
  filter(has_full_submission) %>%
  #mutate(submission_year = submission_year + 2012) %>%
  filter(submission_year <= 2016 & !is.na(ca_continent)) %>%
  group_by(submission_year, ca_continent) %>%
  summarize(
    count = n()
  ) %>%
  ggplot(aes(x = submission_year, y = count, color = ca_continent, shape = ca_continent)) +
  geom_line() +
  geom_point() +
  theme_dakota() +
  labs(x = "Year",
       y = "# Submissions")

country_over_time
```

### Save the file
```{r}
## Save the file
name <- paste0("~/Dropbox/eLife/elife-analysis/figures/supp/", figure_name, figure_extension)
ggsave(name, plot = country_over_time, height = 6, width = 10)
```