### 2019-03-28 figures/gatekeeper_author_outcomes/gatekeeper_author_outcomes.Rmd
updates to figures and labels—prepreparing for re-submission

---
title: "Submission outcomes by author and gatekeeper characteristics"
author: "Dakota Murray"
output: html_notebook
---

### Set parameters for the name and extension of the figure
```{r}
figure_name <- "fig_5"
figure_extension <- ".pdf"

path_to_data <- "https://www.dropbox.com/s/u5lmsyri359mvr9/formatted_elife.csv?dl=1"
```

```{r message=FALSE, warning=FALSE}
# Load required packages
library(tidyverse)
library(grid)
library(gridExtra)
library(stargazer)

# setup theme
theme_dakota <- function () { 
    theme_bw(base_size=10, base_family="Times") %+replace% 
        theme(
            strip.background = element_blank(),
            #strip.text = element_text(face = "bold"),
            plot.title = element_text(size = 20, face = "bold"),
            axis.text.x = element_text(size = 9, margin = margin(t = 5, b = 5), face = "bold"),
            legend.title = element_blank(),
            strip.text = element_text(size = 11),
            legend.position = "bottom"
        )
}

# converts a provided p-value into a representation that is easier to visualize
sig2ast <- function(p) { ifelse(p <= 0.0001, "***", 
                                ifelse(p <= 0.001, "***", 
                                       ifelse(p <= 0.01, "**", 
                                              ifelse(p <= 0.05, "*", 
                                                     ifelse(p < 0.1, "-", "ns")
                                              )
                                       )
                                )
                          )
}

# I copied the below function from a stack overflow question. It should capitalize the first letter of each work in a string
# https://stackoverflow.com/questions/6364783/capitalize-the-first-letter-of-both-words-in-a-two-word-string
simpleCap <- function(x) {
  s <- strsplit(x, " ")[[1]]
  paste(toupper(substring(s, 1,1)), substring(s, 2),
      sep="", collapse=" ")
}
```

### Load the required data
```{r message=FALSE, warning=FALSE}
elife <- read_csv(path_to_data)
head(elife)
```

## Outcomes by author gender and gender composition of authors

> Figure 5. Percentage of full submissions that were accepted, shown by the gender of the last author, and divided by the gender composition of the peer reviewers. Text at the base of each bar indicate the number full submissions within each category of reviewer team and authorship gender. Vertical error bars indicate 95th percentile confidence intervals of the proportion of accepted full submissions. For the composition category of “All women reviewers”, error bars extend beyond scale of figure; see Figure SI.3 to see the error interval. Asterisks indicate significance level of X2 tests of independence on frequency of acceptance by gender of author given each team composition; “ns” indicates no observed statistical significance. “\*“ = p < 0.05; “ns” = p > 0.05.

```{r message=FALSE, warning=FALSE}
sig_list <- list()
for(comp in c("All Men", "All Women", "Mixed")) {
  for(gender_var in c("ca_gender", "fa_gender", "la_gender")) {
    if (!is.na(comp)) {
      sub <- subset(elife, initial_decision_made & full_decision_made & composition == comp)[, c("accepted", gender_var)]
      t <- t(table(sub))[1:2,]
      test = prop.test(t)
      test$n = dim(sub)[1]
      sig_list[length(sig_list) + 1] <- list(test)
    }
  }
}

sig_df <- data.frame(composition = c(rep("All Men", 3), rep("All Women", 3), rep("Mixed", 3)),
                     author_key = rep(c("ca_gender", "fa_gender", "la_gender"), 3),
                     sig = sapply(sig_list, function(test) { sig2ast(test$p.value)})
)

data_homophily <- elife %>%
  filter(initial_decision_made & full_decision_made & !is.na(composition)) %>%
  filter(composition != "Uncertain") %>%
  mutate(total_submissions = n()) %>%
  gather(author_key, value = gender, fa_gender, la_gender) %>%
  filter(gender %in% c("M", "F")) %>%
  group_by(composition, author_key, gender) %>%
  summarize(
    paper_count = n(),
    accept_count = sum(accepted == TRUE),
    acceptance_rate = sum(accepted == TRUE) / paper_count,
    standard_error = 1.96 * sqrt( (acceptance_rate * (1 - acceptance_rate)) / paper_count),
    acceptance_rate = acceptance_rate * 100,
    standard_error = standard_error * 100,
    lower = acceptance_rate - standard_error,
    upper = acceptance_rate + standard_error
  ) %>%
  ungroup() %>%
  left_join(sig_df, by = c("composition", "author_key")) %>%
  mutate(
    author_key = factor(author_key,
                        levels = c("ca_gender", "fa_gender", "la_gender"),
                        labels = c("ca_gender" = "Corresponding Author", "fa_gender" = "First Author", "la_gender" = "Last Author"))
    ,
    gender = ifelse(gender == "M", "Male", "Female"),
    composition = factor(composition, labels = c("All Men" = "All-male", "All Women" = "All-female", "Mixed" = "Mixed-gender"))
  ) %>%
  filter(composition != "Uncertain")

# Construct the final plot
p1 <-  data_homophily %>% 
  filter(author_key == "First Author") %>%
  ggplot(aes(x = gender, y = acceptance_rate, fill = gender)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") +
  geom_text(aes(y = 5, label = paper_count), size = 4, color = "black") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, color = "black", alpha = 0.8) +
  geom_text(aes(x = 1.5, y = 92, label = sig), size = 4, color = "black") +
  geom_segment(aes(x = 1.1, xend = 1.9, y = ifelse(sig != "N/A", 85, 101), yend = ifelse(sig != "N/A", 85, 101)), color = "black") +
  #facet_grid(author_key~composition) +
  facet_wrap(~composition) +
  coord_cartesian(ylim = c(0.0, 100)) +
  theme_dakota() +
  guides(fill = F) +
  labs(y = "% of full submissions accepted",
       x = "First Author Gender",
       title = "A"
  ) +
  theme(axis.title.x = element_text(face = "bold", size = 12),
        axis.title.y = element_text(size = 12),
        strip.text.x = element_text(size = 12),
        strip.text.y = element_text(size = 14, face = "bold"))

# Construct the final plot
p2 <- data_homophily %>% 
  filter(author_key == "Last Author") %>%
  ggplot(aes(x = gender, y = acceptance_rate, fill = gender)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") +
  geom_text(aes(y = 5, label = paper_count), size = 4, color = "black") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, color = "black", alpha = 0.8) +
  geom_text(aes(x = 1.5, y = 92, label = sig), size = 4, color = "black") +
  geom_segment(aes(x = 1.1, xend = 1.9, y = ifelse(sig != "N/A", 85, 101), yend = ifelse(sig != "N/A", 85, 101)), color = "black") +
  #facet_grid(author_key~composition) +
  facet_wrap(~composition) +
  coord_cartesian(ylim = c(0.0, 100)) +
  theme_dakota() +
  guides(fill = F) +
  labs(y = "% of full submissions accepted",
       x = "Last Author Gender",
       title = "B"
  ) +
  theme(axis.title.x = element_text(face = "bold", size = 12),
        axis.title.y = element_text(size = 12),
        strip.text.x = element_text(size = 12),
        strip.text.y = element_text(size = 14, face = "bold"))

gender_homophily <- grid.arrange(p1, p2, nrow = 2)
```

###Signif. test resutls for figure 5
```{r}
names(sig_list) <- c("Reviewers all Men - Corresponding", "Reviewers all Men - First", "Reviewers all Men - Last", 
                     "Reviewers all Women - Corresponding", "Reviewers all Women - First", "Reviewers all Women - Last", 
                     "Mixed Gender Reviewers - Corresponding", "Mixed Gender Reviewers - First", "Mixed Gender Reviewers - Last")

fig4_sig_table <- data.frame(
  composition = sapply(strsplit(names(sig_list), " - "), function(x) x[1]),
  authorship = sapply(strsplit(names(sig_list), " - "), function(x) x[2]),
  #n = sapply(sig_list, function(x) sum(x$n)),
  #df = sapply(sig_list, function(x) x$parameter),
  estimate = sapply(sig_list, function(x) abs(x$estimate[2] - x$estimate[1])),
  statistic = sapply(sig_list, function(x) x$statistic),
  lower = sapply(sig_list, function(x) x$conf.int[1]),
  upper = sapply(sig_list, function(x) x$conf.int[2]),
  #conf = sapply(sig_list, function(x) x$conf.int),
  p.value = round(sapply(sig_list, function(x) x$p.value), 4)
)

rownames(fig4_sig_table) <- NULL
fig4_sig_table
```

###Data used to produce figure 5
```{r}
data_homophily %>%
  select(-standard_error, -lower, -upper, -sig, -author_key)
```

### Number of full submissions
```{r}
full <- elife %>% filter(has_full_submission)
dim(full)[1]
```

###Number of full submissions for which a final decision of accept or reject was made
```{r}
table(full$final_decision_made)
```

###Reviewer team composition for full submissions with a final decision
```{r}
full_final <- full %>% filter(final_decision_made)
table(full_final$composition)
```

As %
```{r}
prop.table(table(full_final$composition))
```

## National composition of author and gatekeepers

> Figure 7. Left: acceptance rate of full submissions compared between presence and absence of homogeneity between the national affiliation of the corresponding author and of at least one. Difference is shown comparing the results for all submissions (top), for all submissions that do not have corresponding authors from the U.S. (middle), and for all submissions that do not have a corresponding author from the U.S., U.K., or Germany (bottom). Right: acceptance rate of full submissions by national homogeneity, shown by individual countries. Included here are the top eight most prolific countries in terms of number of initial submissions. For both panels: vertical error bars indicate 95th percentile confidence intervals for the proportion of accepted full submissions. Values at the base of each bar indicate the number of observations within that combination of country and homophily variables. Asterisks indicate significance level of X2 tests of independence comparing frequency of accepted full submissions between presence and absence of homophily and within each country. “\*\*\*\*” = p < 0.0001; “\*\*\*” = p < 0.01”; “\*“ = p < 0.05; “.” = p < 0.1; “ns” = p > 0.05.

Note: this figure is constructed in multiple parts

```{r message=FALSE, warning=FALSE}
countries <- c("Canada", "China", "France", "Germany", "Japan", "Switzerland", "United Kingdom", "United States")
sig_list <- list()

for (country in countries) {
  print(country)
  s <- subset(elife, full_decision_made == T & has_full_submission & ca_country == country)
  t <- table(s$la_country_homophily, s$accepted)
  test = prop.test(t[, 1:2])
  sig_list[length(sig_list) + 1] <- list(country = test)
}

sig_df <- data.frame(
  ca_country = countries,
  sig = sapply(sig_list, function(l) {sig2ast(l$p.value) })
)

countries_separate <- elife %>%
  filter(full_decision_made & has_full_submission & ca_country %in% countries) %>%
  group_by(ca_country, la_country_homophily) %>%
  summarize(
    paper_count = n(),
    accept_count = sum(accepted == TRUE, na.rm = T),
    encourage_accept_rate = accept_count / paper_count,
    se = 1.96 * sqrt( (encourage_accept_rate * (1 - encourage_accept_rate)) / paper_count),
    encourage_accept_rate = encourage_accept_rate * 100,
    se = se * 100,
    lower = encourage_accept_rate - se,
    upper = encourage_accept_rate + se,
    upper = ifelse(upper > 100, 100, upper)
  ) %>%
  group_by(ca_country) %>%
  mutate(totals = sum(paper_count)) %>%
  ungroup() %>%
  left_join(sig_df) %>%
  #top_n(16, totals) %>% # 8 countries, 3 authorship types per country, therefore first 24 rows
  mutate(ca_country = factor(sapply(as.character(ca_country), simpleCap))) %>%
  # Start ggplot
  ggplot(aes(x = la_country_homophily, y = encourage_accept_rate, fill = la_country_homophily)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black", alpha = 0.8) +
  geom_errorbar(aes(ymin = lower, ymax = upper), color = "black", width = 0.2) +
  geom_text(aes(label = paper_count, y = 7.5), size = 4) +
  geom_text(aes(label = sig, x = 1.5, y = 86), size = 4) +
  geom_segment(aes(x = 1, xend = 2, y = ifelse(sig != "N/A", 80, 101), yend = ifelse(sig != "N/A", 80, 101)), color = "black") +
  facet_wrap(~ca_country, nrow = 2, ncol = 4, labeller = label_wrap_gen()) +
  theme_dakota() +
  guides(fill = FALSE) +
  scale_fill_brewer(palette = "Oranges", labels = c("No Homogeneity", "Homogeneity")) +
  scale_y_continuous(limits = c(0, 100), position = "right") +
  scale_x_discrete(labels = c("No Homogeneity", "Homogeneity")) +
  labs(
    x = "",
    y = "% of full submissions accepted",
    title = "D"
  ) +
  theme(
    axis.text.x = element_text(size = 11, angle = 330, vjust = 1, hjust = 0, face = "bold"),
    axis.title.y = element_text(size = 12)
  )

countries_separate 
```

And here are the results from the tests used in the graph above
```{r}
fig7_sig_table <- data.frame(
  country = countries,
  p.value = round(sapply(sig_list, function(x) x$p.value), 4),
  #n = sapply(sig_list, function(x) (x$n)),
  estimate = sapply(sig_list, function(x) abs(x$estimate[1] - x$estimate[2])),
  statistic = sapply(sig_list, function(x) x$statistic),
  lower = sapply(sig_list, function(x) x$conf.int[1]),
  upper = sapply(sig_list, function(x) x$conf.int[2])
)
rownames(fig7_sig_table) <- NULL
fig7_sig_table
```

```{r message=FALSE, warning=FALSE}
# Get the data, all countries included
d1 <- elife %>%
  filter(full_decision_made & has_full_submission)

# exclude United States
d2 <- elife %>%
  filter(full_decision_made & has_full_submission) %>%
  filter(ca_country != "United States")

# Excluding all founding countries: U.S., U.K., and Germany
d2.5 <- elife %>%
  filter(full_decision_made & has_full_submission) %>%
  filter(!ca_country %in% c("United States", "United Kingdom"))

# Excluding all founding countries: U.S., U.K., and Germany
d3 <- elife %>%
  filter(full_decision_made & has_full_submission) %>%
  filter(!ca_country %in% c("United States", "United Kingdom", "Germany"))

# we also need to get the statistical test infomration for each
t1 <- prop.test(table(d1$la_country_homophily, d1$accepted))
t2 <- prop.test(table(d2$la_country_homophily, d2$accepted))
t2.5 <- prop.test(table(d2.5$la_country_homophily, d2.5$accepted))
t3 <- prop.test(table(d3$la_country_homophily, d3$accepted))

sig_list <- list(t1, t2, t2.5, t3)

sig_df <- data.frame(
  type = c("all", "no us", "no usuk", "no founder"),
  sig = sapply(sig_list, function(l) {sig2ast(l$p.value) })
)

d1.summ <- d1 %>%
  group_by(la_country_homophily) %>%
  summarize(
    paper_count = n(),
    accept_count = sum(accepted == TRUE, na.rm = T),
    encourage_accept_rate = accept_count / paper_count,
    se = 1.96 * sqrt( (encourage_accept_rate * (1 - encourage_accept_rate)) / paper_count),
    encourage_accept_rate = encourage_accept_rate * 100,
    se = se * 100,
    lower = encourage_accept_rate - se,
    upper = encourage_accept_rate + se,
    upper = ifelse(upper > 100, 100, upper),
    type = "all"
  )

d2.summ <- d2 %>%
  group_by(la_country_homophily) %>%
  summarize(
    paper_count = n(),
    accept_count = sum(accepted == TRUE, na.rm = T),
    encourage_accept_rate = accept_count / paper_count,
    se = 1.96 * sqrt( (encourage_accept_rate * (1 - encourage_accept_rate)) / paper_count),
    encourage_accept_rate = encourage_accept_rate * 100,
    se = se * 100,
    lower = encourage_accept_rate - se,
    upper = encourage_accept_rate + se,
    upper = ifelse(upper > 100, 100, upper),
    type = "no us"
  )

d25.summ <- d2.5 %>%
  group_by(la_country_homophily) %>%
  summarize(
    paper_count = n(),
    accept_count = sum(accepted == TRUE, na.rm = T),
    encourage_accept_rate = accept_count / paper_count,
    se = 1.96 * sqrt( (encourage_accept_rate * (1 - encourage_accept_rate)) / paper_count),
    encourage_accept_rate = encourage_accept_rate * 100,
    se = se * 100,
    lower = encourage_accept_rate - se,
    upper = encourage_accept_rate + se,
    upper = ifelse(upper > 100, 100, upper),
    type = "no usuk"
  )

d3.summ <- d3 %>%
  group_by(la_country_homophily) %>%
  summarize(
    paper_count = n(),
    accept_count = sum(accepted == TRUE, na.rm = T),
    encourage_accept_rate = accept_count / paper_count,
    se = 1.96 * sqrt( (encourage_accept_rate * (1 - encourage_accept_rate)) / paper_count),
    encourage_accept_rate = encourage_accept_rate * 100,
    se = se * 100,
    lower = encourage_accept_rate - se,
    upper = encourage_accept_rate + se,
    upper = ifelse(upper > 100, 100, upper),
    type = "no founder"
  )

countries_pooled <- d1.summ %>%
  union(d2.summ) %>%
  union(d25.summ) %>%
  #union(d3.summ) %>%
  left_join(sig_df, by = "type") %>%
  mutate(type = factor(type, 
                       levels = c("all", "no us", "no usuk"),
                       labels = c("All Countries", "No U.S.", "No U.S., U.K.")
                       )
         ) %>%
  ggplot(aes(x = la_country_homophily, y = encourage_accept_rate, fill = la_country_homophily)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black", alpha = 0.8) +
  geom_errorbar(aes(ymin = lower, ymax = upper), color = "black", width = 0.2) +
  geom_text(aes(label = paper_count, y = 10), size = 4) +
  geom_text(aes(label = sig, x = 1.5, y = 85), size = 4) +
  geom_segment(aes(x = 1, xend = 2, y = ifelse(sig != "N/A", 80, 101), yend = ifelse(sig != "N/A", 80, 101)), color = "black") +
  facet_wrap(~type, ncol = 4, labeller = label_wrap_gen(20)) +
  coord_cartesian(ylim = c(0.0, 100)) +
  theme_dakota() +
  guides(fill = FALSE) +
  ylim(0, 1) +
  scale_fill_brewer(palette = "Oranges", labels = c("No Homogeneity", "Homogeneity")) +
  scale_x_discrete(labels = c("No Homogeneity", "Homogeneity")) +
  scale_y_continuous(position = "right") +
  labs(
    x = "",
    y = "% of full submissions accepted",
    title = "C"
  ) +
  theme(axis.text.x = element_text(size = 11, angle = 330, vjust = 1, hjust = 0, face = "bold"),
        axis.title.y = element_text(size = 12))

countries_pooled
```

### Testing the pooled country proportions
```{r}
fig7.B_sig_table <- data.frame(
  type  = c("all", "no us", "no usuk", "no founer"),
  p.value = round(sapply(sig_list, function(x) x$p.value), 4),
  #n = sapply(sig_list, function(x) sum(x$observed)),
  estimate = sapply(sig_list, function(x) abs(x$estimate[1] - x$estimate[2])),
  statistic = sapply(sig_list, function(x) x$statistic),
  df = sapply(sig_list, function(x) x$parameter),
  lower = sapply(sig_list, function(x) x$conf.int[1]),
  upper = sapply(sig_list, function(x) x$conf.int[2])
)

rownames(fig7.B_sig_table) <- NULL
fig7.B_sig_table
```

### Presence of author-reviewer national homogeny by country of correposnding author
```{r rows.print = 20}
full <- elife %>% 
  filter(has_full_submission) %>%
  group_by(ca_country) %>%
  summarize(
    submissions = n(),
    prop_homogeny = sum(la_country_homophily, na.rm = T) / submissions
  ) %>%
  arrange(-prop_homogeny) %>%
  top_n(20) %>%
  mutate(prop_homogeny = round(prop_homogeny, 3))

names(full) <- c("Country", "# Submissions", "% Country Homogeneity")

stargazer(full, type = "latex", summary = FALSE, rownames = FALSE, font.size = "scriptsize", digits = )
```

### Proportion of manuscripts where the corresponding author country is the same as the first and last author country
```{r}
prop.table(table(elife$ca_country == elife$la_country))
prop.table(table(elife$ca_country == elife$la_country))
```

```{r}
us_submissions <- elife %>%
  filter(has_full_submission) %>%
  filter(ca_country == "United States")

prop.table(table(us_submissions$la_country_homophily))
```

```{r}
uk_submissions <- elife %>%
  filter(has_full_submission) %>%
  filter(ca_country == "United Kingdom")

prop.table(table(uk_submissions$la_country_homophily))
```

```{r}
germ_submissions <- elife %>%
  filter(has_full_submission) %>%
  filter(ca_country == "Germany")

prop.table(table(germ_submissions$la_country_homophily))
```

```{r}
submissions <- elife %>%
  filter(has_full_submission) %>%
  filter(ca_country == "Japan")

prop.table(table(submissions$la_country_homophily))
```

```{r}
submissions <- elife %>%
  filter(has_full_submission) %>%
  filter(ca_country == "China")

prop.table(table(submissions$la_country_homophily))
```

```{r}
with_homogeny <- elife %>%
  filter(la_country_homophily)

table(with_homogeny$ca_country)
prop.table(table(with_homogeny$ca_country))
```

```{r}
g1 <- grid.arrange(countries_pooled, countries_separate, ncol = 1, heights = 2:3)
g <- grid.arrange(gender_homophily, g1, ncol = 2)
```

```{r}
## Save the file
name <- paste0("~/Dropbox/eLife/elife-analysis/figures/main/", figure_name, figure_extension)
ggsave(name, plot = g, height = 8, width = 10)
```

```{r}
## Save the file
name <- paste0("~/Dropbox/eLife/elife-analysis/figures/main/", "si_x", figure_extension)
ggsave(name, plot = countries_pooled, height = 5, width = 6)

```