### 2019-07-15 figures/regression_analysis/regression_analysis_simple.rmd

---
title: "Regression Analysis—Simple"
author: "Dakota Murray"
output: html_notebook
---

### Set parameters for the name and extension of the figure
```{r}
figure_name <- "fig_6"
figure_extension <- ".pdf"

path_to_data <- "https://www.dropbox.com/s/97im2cayb1trlax/formatted_elife.csv?dl=1"
```

```{r message=FALSE, warning=FALSE}
# Load required packages
library(tidyverse)
library(lmtest)
library(tidyr)
library(stargazer)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(grid)
library(gridExtra)
library(broom)

# converts a provided p-value into a representation that is easier to visualize
sig2ast <- function(p) { ifelse(p <= 0.0001, "***", 
                                ifelse(p <= 0.001, "***", 
                                       ifelse(p <= 0.01, "**", 
                                              ifelse(p <= 0.05, "*", 
                                                     ifelse(p < 0.1, ".", "ns")
                                              )
                                       )
                                )
                          )
}
```

### Load the required data
```{r message=FALSE, warning=FALSE}
# Replace this string 
elife <- read.csv(path_to_data)

elife$ca_continent <- relevel(elife$ca_continent, ref = "North America")
elife$fa_continent <- relevel(elife$fa_continent, ref = "North America")
elife$la_continent <- relevel(elife$la_continent, ref = "North America")
elife$se_continent <- relevel(elife$se_continent, ref = "North America")

elife$composition <- relevel(elife$composition, ref = "Mixed")

elife$la_institution_rank <- with(elife, ifelse(la_institution_rank == "Top", "Top", "Sandard/Unranked/Unknown"))
elife$ca_institution_rank <- with(elife, ifelse(ca_institution_rank == "Top", "Top", "Sandard/Unranked/Unknown"))

elife$submission_year <- elife$submission_year - 2012
dim(elife)
```

### Define initial and full submissions
```{r}
# First, filter the data by removing the observations from Antarctica 
# add a variable for "included"
initial_submissions <- elife %>% filter(initial_decision_made)
initial_submissions$included <- with(initial_submissions, initial_decision_made & ca_continent != "Antarctica" & !is.na(ca_institution_rank) & !is.na(ca_gender) & !is.na(ca_se_dist))

full_submissions <- elife %>% filter(full_decision_made)
full_submissions$included <- with(full_submissions, la_continent != "Antarctica" & fa_continent != "Antarctica" & !is.na(la_gender) & la_gender != "UNK" & composition != "Uncertain" & !is.na(la_reviewer_dist) & !is.na(la_reviewer_dist_zero) & !is.na(la_institution_rank) & !is.na(la_continent) & !is.na(fa_gender))
```

### Initial submissions model
```{r}
initial_model <- glm(encouraged ~ submission_year + submission_type + ca_gender + ca_institution_rank + ca_continent, data = subset(initial_submissions, included), family = "binomial")
summary(initial_model)
```

```{r}
initial.cov.labels = c(
               "submission_year" = "Submission Year", 
               "submission_typeSR" = "Submission Type = SR", 
               "submission_typeTR" = "Submission Type = TR", 
               "ca_genderM" = "Corr. Author is Male",
               "ca_genderUNK" = "Corr. Author is Unknown Gender",
               "ca_institution_rankTop" = "Corr. Author Inst. Top",
               "ca_continentAfrica" = "Corr. Author from Africa",
               "ca_continentAsia" = "Corr. Author from Asia",
               "ca_continentEurope" = "Corr. Author from Europe",
               "ca_continentOceania" = "Corr. Author from Oceania",
               "ca_continentSouth America" = "Corr. Author from South America"
               )

stargazer(initial_model, 
          ci = T, 
          align = T,
          t.auto = F,
          p.auto = F,
          apply.coef = exp,
          covariate.labels = initial.cov.labels, 
          model.names = T,
          style = "ajs",
          type = "latex"
          )

#cat(paste(out, collapse = "\n"), "\n", file="~/Desktop/encourage_model.txt", append=F)
```

### Full submissions model
```{r}
full_model <- glm(accepted ~ submission_year + submission_type + fa_gender + la_gender + la_institution_rank + la_continent, data = subset(full_submissions, included), family = "binomial")
summary(full_model)
```

### Construct a table for the regression model for full submissions
```{r}
full.cov.labels = c("submission_year" = "Submission Year", 
               "submission_typeSR" = "Submission Type = SR", 
               "submission_typeTR" = "Submission Type = TR", 
               "fa_genderM" = "First Author is Male",
               "fa_genderUNK" = "First Author is Unknown Gender",
               "la_genderM" = "Last Author is Male",
               "la_institution_rankTop" = "Last Author Inst. Top",
               "la_continentAfrica" = "Last Author from Africa",
               "la_continentAsia" = "Last Author from Asia",
               "la_continentEurope" = "Last Author from Europe",
               "la_continentOceania" = "Last Author from Oceania",
               "la_continentSouth America" = "Last Author from South America"
               )

stargazer(full_model, 
          ci = T, 
          align = T, 
          t.auto = F,
          p.auto = F,
          apply.coef = exp,
          covariate.labels = full.cov.labels, 
          model.names = T,
          style = "ajs",
          font.size = "scriptsize",
          type = "latex"
          )

#out
#cat(paste(out, collapse = "\n"), "\n", file="~/Desktop/accept_model.txt", append=TRUE)
```

```{r}
initial_tidy <- tidy(initial_model, conf.int = T, exponentiate = T) %>%
  mutate(type = "Initial Submissions")
full_tidy <- tidy(full_model, conf.int = T, exponentiate = T) %>%
  mutate(type = "Full Submissions")

all_submissions <- rbind(initial_tidy, full_tidy)
```

```{r}
initial.plot <- initial_tidy %>%
  filter(term != "(Intercept)") %>%
  mutate(
    #type = factor(type, levels = c("Main Effects", "Global Interactions")),
    ast = sig2ast(p.value),
    ast.factor = factor(ast, 
                 levels = c("ns", ".", "*", "**", "***"), 
                 labels = c("p > 0.1", "p < 0.1", "p < 0.05", "p < 0.01", "p < 0.001")),
    color.sig = ifelse(p.value < 0.1, ifelse(estimate > 1, "Positive", "Negative"), "Not Signigicant"),
    term = as.factor(term),
    term = plyr::revalue(term, initial.cov.labels),
    term = reorder(term, estimate)
  ) %>%
  ggplot(aes(x = term, y = estimate, color = color.sig, shape = ast.factor)) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  coord_flip() +
  #facet_grid(~type, scales = "free") +
  geom_text(aes(label = paste0(round(estimate, 3), 
                               ifelse(ast %in% c("ns", "."), "", ast)), 
                x = term, y = estimate), vjust = -0.6, hjust = -0.1, show.legend = F) +
  scale_color_manual(name = "Direction of Effect", values = c("firebrick", "grey", "blue")) +
  scale_shape_manual(name = "Significance Level", values = c(1, 10, 16, 17, 18), drop = F) +
  geom_hline(yintercept = 1, color = "black") +
  coord_flip(ylim = c(0.1, 3)) +
  theme_minimal() +
  guides(color = guide_legend(title.position = "top"),
         shape = guide_legend(title.position = "top")) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(size = 9),
        legend.position = "bottom",
        plot.title = element_text(size = 12, face = "bold")) +
  labs(y = "Odds Ratio",
       title = "A. Initial Submissions")

initial.plot 
```

```{r}
full.plot <- full_tidy %>%
  filter(term != "(Intercept)") %>%
  mutate(
    ast = sig2ast(p.value),
    ast.factor = factor(ast, 
                 levels = c("ns", ".", "*", "**", "***"), 
                 labels = c("p > 0.1", "p < 0.1", "p < 0.05", "p < 0.01", "p < 0.001")),
    color.sig = ifelse(p.value < 0.1, ifelse(estimate > 1, "Positive", "Negative"), "Not Signigicant"),
    term = as.factor(term),
    term = plyr::revalue(term, full.cov.labels),
    term = reorder(term, estimate)
  ) %>%
  ggplot(aes(x = term, y = estimate, color = color.sig, shape = ast.factor)) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  #coord_flip() +
  #facet_grid(~type, scales = "free") +
  geom_text(aes(label = paste0(round(estimate, 3), 
                               ifelse(ast %in% c("ns", "."), "", ast)), 
                x = term, y = estimate), vjust = -0.6, hjust = -0.1, show.legend = F) +
  scale_color_manual(name = "Direction of Effect", values = c("firebrick", "grey", "blue")) +
  scale_shape_manual(name = "Significance Level", values = c(1, 10, 16, 15, 18), drop = F) +
  geom_hline(yintercept = 1, color = "black") +
  coord_flip(ylim = c(0.1, 3)) +
  theme_minimal() +
  guides(color = guide_legend(title.position = "top"),
         shape = guide_legend(title.position = "top")) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_text(size = 9),
        legend.position = "bottom",
        plot.title = element_text(size = 12, face = "bold")) +
  labs(y = "Odds Ratio",
       title = "B. Full Submissions")

full.plot
```

```{r}
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

mylegend <- g_legend(initial.plot)

both.plots <- grid.arrange(arrangeGrob(initial.plot + theme(legend.position="none"),
                         full.plot + theme(legend.position="none"),
                         nrow=1),
             mylegend, nrow=2,heights=c(10, 2))

both.plots
```

### Save the output
```{r}
name <- paste0("~/Dropbox/eLife/elife-analysis/figures/main/", figure_name, figure_extension)
ggsave(name, plot = both.plots, height = 8, width = 8)
```