### 2019-07-30 figures/general_information/supp_gender_prop_by_country.rmd

---
title: "Country and Gender Demographics"
author: "Dakota Murray"
output: html_notebook
---

### Set parameters for the name and extension of the figure
```{r}
figure_name <- "fig_si5"
figure_extension <- ".pdf"

path_to_data <- "https://www.dropbox.com/s/97im2cayb1trlax/formatted_elife.csv?dl=1"
```

```{r message=FALSE, warning=FALSE}
# Load required packages
library(tidyverse)
library(grid)
library(gridExtra)

# setup theme
theme_dakota <- function () { 
    theme_bw(base_size=10, base_family="Times") %+replace% 
        theme(
            strip.background = element_blank(),
            strip.text = element_text(size = 12, face = "bold"),
            plot.title = element_text(size = 12, face = "bold"),
            axis.text.x = element_text(size = 12, margin = margin(t = 5, b = 5)),
            axis.title.x = element_text(size = 12, face = "bold"),
            axis.title.y = element_text(size = 12, angle = 90),
            axis.text.y = element_text(size = 10, margin = margin(t = 5, b = 5), hjust = 1),
            legend.title = element_blank(),
            legend.position = "bottom"
        )
}
```

### Load the required data
```{r message=FALSE, warning=FALSE}
elife <- read_csv(path_to_data)

#full_submissions <- elife %>% filter(full_decision_made)
#full_submissions$included <- with(full_submissions, la_continent != "Antarctica" & fa_continent != "Antarctica" & !is.na(la_gender) & la_gender != "UNK" & composition != "Uncertain" & !is.na(la_reviewer_dist) & !is.na(la_reviewer_dist_zero) & !is.na(la_institution_rank) & !is.na(la_continent) & !is.na(fa_gender))

head(full_submissions)
```

### 
```{r}
gender_by_country <- elife %>%
  mutate(ca_country = factor(sapply(as.character(ca_country), simpleCap))) %>% # capitalize first letter of country stirngs
  group_by(ca_country) %>%
  mutate(total = n()) %>%
  filter(total > 200) %>%
  filter(ca_gender == "F") %>%
  group_by(ca_country) %>%
  summarize(
    prop = n() / total[1]
  ) %>%
  mutate(
    ca_country = reorder(ca_country, -prop)
  ) %>%
  ggplot(aes(x = ca_country, y = prop)) +
  geom_bar(stat = "identity", alpha = 0.8, fill = "orange") +
  ylim(0, 0.5) +
  coord_flip() +
  theme_dakota() +
  labs(x = "Country of corresponding author",
       y = "Proportion of initial submissions with a female correpsonding author") +
  scale_fill_discrete(labels = c("F" = "Female", "M" = "Male"))

gender_by_country
```

### Save the file
```{r}
name <- paste0("~/Dropbox/eLife/elife-analysis/figures/supp/", figure_name, figure_extension)
ggsave(name, plot = gender_by_country, height = 4, width = 7)
```