### 2019-08-02 figures/regression_analysis/regression_analysis_interaction.rmd
added final tweaks to figures, largely just swapping output names and tweaking labels

---
title: "Regression Analysis—Interactions"
author: "Dakota Murray"
output: html_notebook
---

### Set parameters for the name and extension of the figure
```{r}
figure_name <- "fig_7"
figure_extension <- ".pdf"

path_to_data <- "https://www.dropbox.com/s/97im2cayb1trlax/formatted_elife.csv?dl=1"
```

```{r message=FALSE, warning=FALSE}
# Load required packages
library(tidyverse)
library(lmtest)
library(tidyr)
library(stargazer)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(grid)
library(gridExtra)
library(broom)

# converts a provided p-value into a representation that is easier to visualize
sig2ast <- function(p) { ifelse(p <= 0.0001, "***", 
                                ifelse(p <= 0.001, "***", 
                                       ifelse(p <= 0.01, "**", 
                                              ifelse(p <= 0.05, "*", 
                                                     ifelse(p < 0.1, ".", "ns")
                                              )
                                       )
                                )
                          )
}
```

### Load the required data
```{r message=FALSE, warning=FALSE}
# Replace this string 
elife <- read.csv(path_to_data)

elife$ca_continent <- relevel(elife$ca_continent, ref = "North America")
elife$fa_continent <- relevel(elife$fa_continent, ref = "North America")
elife$la_continent <- relevel(elife$la_continent, ref = "North America")
elife$se_continent <- relevel(elife$se_continent, ref = "North America")

elife$composition <- relevel(elife$composition, ref = "Mixed")

elife$la_institution_rank <- with(elife, ifelse(la_institution_rank == "Top", "Top", "Sandard/Unranked/Unknown"))
elife$ca_institution_rank <- with(elife, ifelse(ca_institution_rank == "Top", "Top", "Sandard/Unranked/Unknown"))

elife$submission_year <- elife$submission_year-2012
dim(elife)
```

### Define initial and full submissions
```{r}
# First, filter the data by removing the observations from Antarctica 
# add a variable for "included"
initial_submissions <- elife %>% filter(initial_decision_made)
initial_submissions$included <- with(initial_submissions, initial_decision_made & ca_continent != "Antarctica" & !is.na(ca_institution_rank) & !is.na(ca_gender) & !is.na(ca_se_dist))

full_submissions <- elife %>% filter(full_decision_made)
full_submissions$included <- with(full_submissions, la_continent != "Antarctica" & fa_continent != "Antarctica" & !is.na(la_gender) & la_gender != "UNK" & composition != "Uncertain" & !is.na(la_reviewer_dist) & !is.na(la_reviewer_dist_zero) & !is.na(la_institution_rank) & !is.na(la_continent) & !is.na(fa_gender))
```

### Construct the initial submissions model
```{r}
initial_model <- glm(encouraged ~ submission_year + submission_type + ca_gender +  ca_institution_rank + ca_continent + I(ca_se_dist / 1000000) + ca_se_dist_zero + ca_se_dist_zero, data = subset(initial_submissions, included), family = "binomial")
summary(initial_model)
```

### Construct a regression table for the 
```{r}
initial.cov.labels = c(
               "submission_year" = "Submission Year", 
               "submission_typeSR" = "Submission Type = SR", 
               "submission_typeTR" = "Submission Type = TR", 
               "ca_genderM" = "Corr. Author is Male",
               "ca_genderUNK" = "Corr. Author Gender UNK",
               "ca_institution_rankTop" = "Corr. Author Inst. Top",
               "ca_continentAfrica" = "Corr. Author from Africa",
               "ca_continentAsia" = "Corr. Author from Asia",
               "ca_continentEurope" = "Corr. Author from Europe",
               "ca_continentOceania" = "Corr. Author from Oceania",
               "ca_continentSouth America" = "Corr. Author from South America",
               "I(ca_se_dist/1e+06)" = "Corr. Author-Editor Geo. Distance",
               "ca_se_dist_zeroTRUE" = "Corr. Author-Editor Geo. Distance = 0"
               )

stargazer(initial_model, 
            ci = T, 
            t.auto = F,
            p.auto = F,
            align = T, 
            apply.coef = exp,
            covariate.labels = initial.cov.labels, 
            model.names = T,
            style = "ajs"
            )

#cat(paste(out, collapse = "\n"), "\n", file="~/Desktop/encourage_model.txt", append=F)
#out
```

### Construct the interaction variable
```{r}
assign_interaction_term = function(la_gender, composition) {
  if (la_gender == "M") {
    if (composition == "All Men") {
      return("Male Author, All Men")
    } else if (composition == "All Women") {
      return("Male Author, All Women")
    } else if (composition == "Mixed") {
      return("Male Author, Mixed")
    } else {
      return(NA)
    }
  } else if (la_gender == "F") {
    if (composition == "All Men") {
      return("Female Author, All Men")
    } else if (composition == "All Women") {
      return("Female Author, All Women")
    } else if (composition == "Mixed") {
      return("Female Author, Mixed")
    } else {
      return(NA)
    }
  } else {
    return(NA)
  }
}

full_with_inter <- full_submissions %>%
  filter(composition != "Uncertain") %>%
  rowwise() %>%
  mutate(
    author_reviewer_interaction = assign_interaction_term(la_gender, composition)
  ) %>%
  ungroup() %>%
  mutate(author_reviewer_interaction = as.factor(author_reviewer_interaction))

table(full_with_inter$author_reviewer_interaction)
```

### Construct the model for full submissions with interactions
```{r}
full_model <- glm(accepted ~ submission_year + submission_type + fa_gender + la_institution_rank + la_continent + I(la_reviewer_dist/ 1000000) + la_reviewer_dist_zero + author_reviewer_interaction, data = subset(full_with_inter, included), family = "binomial")
summary(full_model)
```

```{r}
main_effects <- glm(accepted ~ submission_year + submission_type + fa_gender + la_institution_rank + la_gender + la_continent + I(la_reviewer_dist/ 1000000) + la_reviewer_dist_zero + composition, data = subset(full_with_inter, included), family = "binomial")
summary(main_effects)
```

```{r}
standard_interaction <- glm(accepted ~ submission_year + submission_type + fa_gender + la_institution_rank + la_gender + la_continent + I(la_reviewer_dist/ 1000000) + la_reviewer_dist_zero + composition + la_gender:composition, data = subset(full_with_inter, included), family = "binomial")
summary(standard_interaction)
```

### Run ANOVA on the model...
```{r}
anova(standard_interaction, test = "Chisq")
```

Mixed-gender only
```{r}
mixed <- glm(accepted ~ submission_year + submission_type + fa_gender +  la_institution_rank + la_gender + la_continent + I(la_reviewer_dist/ 1000000) + la_reviewer_dist_zero, data = subset(full_with_inter, included & composition == "Mixed"), family = "binomial")
summary(mixed)
```

All-Male
```{r}
all_male <- glm(accepted ~ submission_year + submission_type + fa_gender + la_institution_rank + la_gender + la_continent + I(la_reviewer_dist/ 1000000) + la_reviewer_dist_zero, data = subset(full_with_inter, included & composition == "All Men"), family = "binomial")
summary(all_male)
```

```{r}
main_effects_tidy <- tidy(main_effects, conf.int = T, exponentiate = T) %>%
  mutate(type = "Main Effects")
standard_interaction_tidy <- tidy(standard_interaction, conf.int = T, exponentiate = T) %>%
  mutate(type = "Standard Interaction")
mixed_only_tidy <- tidy(mixed, conf.int = T, exponentiate = T) %>%
  mutate(type = "Mixed Reviewers Only")
all_male_only_tidy <- tidy(all_male, conf.int = T, exponentiate = T) %>%
  mutate(type = "All-Male Reviewers Only")
global_tidy <- tidy(full_model, conf.int = T, exponentiate = T) %>%
  mutate(type = "Gender homogeny interaction")

all_models <- rbind(main_effects_tidy,
      standard_interaction_tidy,
      mixed_only_tidy,
      all_male_only_tidy,
      global_tidy)
```

```{r}
full.cov.labels = c(
               "submission_year" = "Submission Year", 
               "submission_typeSR" = "Submission Type = SR", 
               "submission_typeTR" = "Submission Type = TR", 
               "fa_genderM" = "First Author is Male",
               "fa_genderUNK" = "First Author is Unknown Gender",
               "compositionAll Men" = "All Male Rev.",
               "compositionAll Women" = "All Female Rev.",
               "la_institution_rankTop" = "Last Author Inst. Top",
               "la_genderM" = "Last Author is Male",
               "la_continentAfrica" = "Last Author from Africa",
               "la_continentAsia" = "Last Author from Asia",
               "la_continentEurope" = "Last Author from Europe",
               "la_continentOceania" = "Last Author from Oceania",
               "la_continentSouth America" = "Last Author from South America",
               "I(la_reviewer_dist/1e+06)" = "Last Author-Reviewers Geo. Distance",
               "la_reviewer_dist_zeroTRUE" = "Last Author-Reviewers Geo. Distance = 0",
               "la_genderM:compositionAll Men" = "Last Author Male*All Male Rev.",
               "la_genderM:compositionAll Women" = "Last Author Male*All Women Rev.",
               "author_reviewer_interactionFemale Author, All Women" = "Last Author Female-All Female Rev.",
               "author_reviewer_interactionFemale Author, Mixed" = "Last Author Female-Mixed Rev.",
               "author_reviewer_interactionMale Author, All Men" = "Last Author Male-All Male Rev.",
               "author_reviewer_interactionMale Author, All Women" = "Last Author Male-All Female Rev.",
               "author_reviewer_interactionMale Author, Mixed" = "Last Author Male-Mixed Rev."
               )

both.full.plots <- all_models %>%
  filter(type == "Gender homogeny interaction" | type == "Main Effects") %>%
  filter(term != "(Intercept)") %>%
  mutate(
    type = factor(type, 
                  levels = c("Main Effects", "Gender homogeny interaction"),
                  labels = c("A.", "B.")),
    ast = sig2ast(p.value),
    ast.factor = factor(ast, 
                 levels = c("ns", ".", "*", "**", "***"), 
                 labels = c("p > 0.1", "p < 0.1", "p < 0.05", "p < 0.01", "p < 0.001")),
    ast.simplified = ifelse(p.value < 0.05, "p < 0.05", "p > 0.05"),
    ast.simplified = factor(ast.simplified, levels = c("p > 0.05", "p < 0.05")),
    color.sig = ifelse(p.value < 0.1, ifelse(estimate > 1, "Positive", "Negative"), "Not Signigicant"),
    term = as.factor(term),
    term = plyr::revalue(term, full.cov.labels),
    term = reorder(term, estimate)
  ) %>%
  ggplot(aes(x = term, y = estimate, color = color.sig, shape = ast.simplified)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  coord_flip(ylim = c(0.1, 3)) +
  facet_grid(~type) +
  geom_text(aes(label = paste0(round(estimate, 3), 
                               ifelse(ast %in% c("ns", "."), "", ast))),
            vjust = -0.6, hjust = -0.1, show.legend = F) +
  scale_color_manual(name = "Direction of Effect", values = c("firebrick", "grey", "blue")) +
  scale_shape_manual(name = "Significance Level", values = c(1, 16)) +
  geom_hline(yintercept = 1, color = "black") +
  theme_minimal() +
  guides(color = guide_legend(title.position = "top"),
         shape = guide_legend(title.position = "top")) +
  theme(axis.title.y = element_blank(),
        strip.text = element_text(size = 20, face = "bold"),
        legend.position = "bottom",
        legend.text = element_text(size = 11),
        legend.title = element_blank(),
        legend.direction = "vertical",
        axis.text.y = element_text(size = 11)) +
  labs(y = "Odds Ratio")

both.full.plots
```

### Construct the plot for full submissions
```{r}
full.cov.labels = c(
               "submission_year" = "Submission Year", 
               "submission_typeSR" = "Submission Type = SR", 
               "submission_typeTR" = "Submission Type = TR", 
               "fa_genderM" = "First Author is Male",
               "fa_genderUNK" = "First Author is Unknown Gender",
               "compositionAll Men" = "All Male Rev.",
               "compositionAll Women" = "All Female Rev.",
               "la_institution_rankTop" = "Last Author Inst. Top",
               "la_genderM" = "Last Author is Male",
               "la_continentAfrica" = "Last Author from Africa",
               "la_continentAsia" = "Last Author from Asia",
               "la_continentEurope" = "Last Author from Europe",
               "la_continentOceania" = "Last Author from Oceania",
               "la_continentSouth America" = "Last Author from South America",
               "I(la_reviewer_dist/1e+06)" = "Last Author-Reviewers Geo. Distance",
               "la_reviewer_dist_zeroTRUE" = "Last Author-Reviewers Geo. Distance = 0",
               "author_reviewer_interactionFemale Author, All Women" = "Last Author Female-All Female Rev.",
               "author_reviewer_interactionFemale Author, Mixed" = "Last Author Female-Mixed Rev.",
               "author_reviewer_interactionMale Author, All Men" = "Last Author Male-All Male Rev.",
               "author_reviewer_interactionMale Author, All Women" = "Last Author Male-All Female Rev.",
               "author_reviewer_interactionMale Author, Mixed" = "Last Author Male-Mixed Rev."
               )

stargazer(full_model, 
          ci = T, 
          align = T, 
          apply.coef = exp,
          t.auto = F,
          p.auto = F,
          covariate.labels = full.cov.labels,
          model.names = T,
          style = "ajs",
          font.size = "scriptsize"
          )
```

```{r}
main.effects.cov.labels = c(
               "submission_year" = "Submission Year", 
               "submission_typeSR" = "Submission Type = SR", 
               "submission_typeTR" = "Submission Type = TR", 
               "fa_genderM" = "First Author is Male",
               "fa_genderUNK" = "First Author is Unknown Gender",
               "la_institution_rankTop" = "Last Author Inst. Top",
               "la_genderM" = "Last Author is Male",
               "la_continentAfrica" = "Last Author from Africa",
               "la_continentAsia" = "Last Author from Asia",
               "la_continentEurope" = "Last Author from Europe",
               "la_continentOceania" = "Last Author from Oceania",
               "la_continentSouth America" = "Last Author from South America",
               "I(la_reviewer_dist/1e+06)" = "Last Author-Reviewers Geo. Distance",
               "la_reviewer_dist_zeroTRUE" = "Last Author-Reviewers Geo. Distance = 0",
               "compositionAll Men" = "All Male Rev.",
               "compositionAll Women" = "All Female Rev."
               )

stargazer(main_effects, 
          ci = T, 
          align = T, 
          apply.coef = exp,
          t.auto = F,
          p.auto = F,
          covariate.labels = main.effects.cov.labels,
          model.names = T,
          style = "ajs",
          font.size = "scriptsize"
          )
```

```{r}
summary(standard_interaction)

```

```{r}

all.cov.labels = c(
               "submission_year" = "Submission Year", 
               "submission_typeSR" = "Submission Type = SR", 
               "submission_typeTR" = "Submission Type = TR", 
               "fa_genderM" = "First Author is Male",
               "fa_genderUNK" = "First Author is Unknown Gender",
               "la_institution_rankTop" = "Last Author Inst. Top",
               "la_genderM" = "Last Author is Male",
               "la_continentAfrica" = "Last Author from Africa",
               "la_continentAsia" = "Last Author from Asia",
               "la_continentEurope" = "Last Author from Europe",
               "la_continentOceania" = "Last Author from Oceania",
               "la_continentSouth America" = "Last Author from South America",
               "I(la_reviewer_dist/1e+06)" = "Last Author-Reviewers Geo. Distance",
               "la_reviewer_dist_zeroTRUE" = "Last Author-Reviewers Geo. Distance = 0",
               "compositionAll Men" = "All Male Rev.",
               "compositionAll Women" = "All Female Rev.",
               "la_genderM:compositionAll Men" = "Last Author Male*All Male Rev.",
               "la_genderM:compositionAll Women" = "Last Author Male*All Women Rev.",
               "author_reviewer_interactionFemale Author, All Women" = "Last Author Female-All Female Rev.",
               "author_reviewer_interactionFemale Author, Mixed" = "Last Author Female-Mixed Rev.",
               "author_reviewer_interactionMale Author, All Men" = "Last Author Male-All Male Rev.",
               "author_reviewer_interactionMale Author, All Women" = "Last Author Male-All Female Rev.",
               "author_reviewer_interactionMale Author, Mixed" = "Last Author Male-Mixed Rev."
               )

stargazer(main_effects, standard_interaction, all_male, mixed, full_model,
          type = "latex",
          style = "ajs",
          ci = T, 
          align = T, 
          apply.coef = exp,
          t.auto = F,
          p.auto = F,
          column.labels = c("Main Effects", "With interaction", "Only All-Male", "Only Mixed", "With Interaction"),
          covariate.labels = all.cov.labels)

```

```{r}
interaction.cov.labels = c(
               "submission_year" = "Submission Year", 
               "submission_typeSR" = "Submission Type = SR", 
               "submission_typeTR" = "Submission Type = TR", 
               "fa_genderM" = "First Author is Male",
               "fa_genderUNK" = "First Author is Unknown Gender",
               "la_institution_rankTop" = "Last Author Inst. Top",
               "la_genderM" = "Last Author Male",
               "la_continentAfrica" = "Last author from Africa",
               "la_continentAsia" = "Last author from Asia",
               "la_continentEurope" = "Last author from Europe",
               "la_continentOceania" = "Last author from Oceania",
               "la_continentSouth America" = "Last author from South America",
               "I(la_reviewer_dist/1e+06)" = "Sum of geo. distance (1000s km)",
               "la_reviewer_dist_zeroTRUE" = "Sum of geo. distance is zero",
               "compositionAll Men" = "All Reviewers Male",
               "compositionAll Women" = "All Reviewers Women",
               "la_genderM:compositionAll Men" = "Last Author Male*All Reviewers Male",
               "la_genderM:compositionAll Women" = "Last Author Male*All Reviewers Female"
               )

stargazer(standard_interaction, 
          ci = T, 
          align = T, 
          apply.coef = exp,
          t.auto = F,
          p.auto = F,
          covariate.labels = interaction.cov.labels,
          model.names = T,
          style = "ajs",
          font.size = "scriptsize"
          )
```

```{r}
interaction.anova.labels = c(
               "submission_year" = "Submission Year", 
               "submission_type" = "Submission Type", 
               "fa_gender" = "First Author Gender",
               "la_institution_rank" = "Last Author Inst. Prestige",
               "la_gender" = "Last Author Gender",
               "la_continent" = "Last author Continent",
               "I(la_reviewer_dist/1e+06)" = "Last Author-Reviewers Geographic Distance",
               "la_reviewer_dist_zero" = "Sum of geo. distance is zero",
               "composition" = "Reviewer Gender Composition",
               "la_gender:composition" = "Last Author Gender*Reviewer Gender Composition"
               )

anova.table <- anova(standard_interaction, test = "Chisq")
anova.tidy <- tidy(anova.table) %>%
  filter(term != "NULL") %>%
  mutate(term = factor(term),
         term = plyr::revalue(term, interaction.anova.labels),
         term = as.character(term),
         Deviance = round(Deviance, 3),
         Resid..Df = round(Resid..Df, 3),
         Resid..Dev = round(Resid..Dev, 3),
         p.value = round(p.value, 5))

anova.tidy

stargazer(anova.tidy,
          summary = FALSE,
          column.labels = c("Term", "df", "Deviance", "Resid. Df", "Resid. Deviance", "P-value"),
          type = "latex",
          style = "ajs",
          font.size = "scriptsize")
```

```{r}
main.effects.cov.labels = c(
               "submission_year" = "Submission Year", 
               "submission_typeSR" = "Submission Type = SR", 
               "submission_typeTR" = "Submission Type = TR", 
               "fa_genderM" = "First Author is Male",
               "fa_genderUNK" = "First Author is Unknown Gender",
               "la_institution_rankTop" = "Last Author Inst. Top",
               "la_genderM" = "Last Author Male",
               "la_continentAfrica" = "Last author from Africa",
               "la_continentAsia" = "Last author from Asia",
               "la_continentEurope" = "Last author from Europe",
               "la_continentOceania" = "Last author from Oceania",
               "la_continentSouth America" = "Last author from South America",
               "I(la_reviewer_dist/1e+06)" = "Sum of geo. distance (1000s km)",
               "la_reviewer_dist_zeroTRUE" = "Sum of geo. distance is zero",
               "compositionAll Men" = "All Reviewers Male",
               "compositionAll Women" = "All Reviewers Women"
               )

stargazer(all_male, mixed, 
          column.labels = c("All Male", "Mixed-Gender"),
          ci = T, 
          align = T, 
          apply.coef = exp,
          t.auto = F,
          p.auto = F,
          covariate.labels = main.effects.cov.labels,
          model.names = T,
          style = "ajs",
          font.size = "scriptsize"
          )

```

### Save the plot
```{r}
name <- paste0("~/Dropbox/eLife/elife-analysis/figures/main/", figure_name, figure_extension)
ggsave(name, plot = both.full.plots, height = 8, width = 8)
```